
```{r}
install.packages("sp")
library(sp)
install.packages("mapproj")
library(mapproj)
install.packages("ggplot2")
library(ggplot2)
install.packages("maptools")
library(maptools)
install.packages("raster")#for bind
library(raster)
install.packages("rgeos")#for gcentroid
library("rgeos")
```

#Loading rds for FINLAND level 2: Maakuntas 
#with anomalies: Uusimaa &Ita Uusimaa divided, three Päijät-Häme divisions.
#Shows 21 regions where the actual number is 18
```{r}
fin<-readRDS("gadm36_FIN_2_sp.rds")
```

```{r}
fin@data$NAME_2
```

#Ita-Uusimaa name to 0
#Extra Paijat Hame set to 0
```{r}
fin@data$NAME_2[8]<-NA
fin@data$NAME_2[10]<-NA
fin@data$NAME_2[17]<-NA
```

#Merging of the Ita-Uusimaa(8) and Uusimma(12) &
#merging of three Paijat Hame regions (3,10,17) by repeating indices
```{r}
fin1<-c( 1 , 2  ,3 , 4 , 5 , 6 , 7, 12  ,9 ,3, 11 ,12 ,13 ,14, 15, 16 ,3 ,18 ,19, 20, 21)
fin.u<-unionSpatialPolygons(fin,fin1) #class SP
```

#A dataframe with merged rows removed 
```{r}
fin.df.1<-fin@data[!is.na(fin@data$NAME_2),]
```

#Converting SP and DF to SPDF 
```{r}
fin.u.spd<-SpatialPolygonsDataFrame(fin.u,fin.df.1)
```

#Loading rds files for ALAND level 0
```{r}
aland<-readRDS("gadm36_ALA_0_sp.rds") #level zero
```

#Merging Finland and Aland SPDFs
```{r}
fin.ala<-bind(fin.u.spd,aland)
```


```{r}
#plot(fin.ala)
```

#Inserting English names
```{r}
fin.ala@data$NAME_2[19]<-"Aland"
fin.ala@data$VARNAME_2[19]<- "Ahvenanmaa|Åland"
```



#Neighbouring nations
```{r}
#Loading rds file for SWEDEN level 0
swed<-readRDS("gadm36_SWE_0_sp.rds")

#Loading rds file for RUSSIA level 0
rus<-readRDS("gadm36_RUS_0_sp.rds")

#Loading rds file for NORWAY level 0
nor<-readRDS("gadm36_NOR_0_sp.rds")

#Loading rds file for ESTONIA level 0
est<-readRDS("gadm36_EST_0_sp.rds")
```


#longitude and latidue for centres
```{r}
centroids<-gCentroid(fin.ala, byid=TRUE)
x0 <- coordinates(centroids)[,1]
y0 <- coordinates(centroids)[,2]
```

#extracting DF from SPDF for ggplot to read - results in dropping of columns with info
```{r}
fin.ala.df <- fortify(fin.ala,region="NAME_2")
```


#Data for population
```{r}
info<-read.csv("maakuntaa_population.csv")
```

#Adding population column
```{r}
fin.ala.df<-merge(fin.ala.df,info,by="id")
```

#Plotting
```{r}
map1<-ggplot(data = fin.ala.df,aes(x = long, y = lat, group=group,fill =pop) )+coord_map(xlim = c(18,33), ylim = c(59,70.5))
map2<-map1 + geom_polygon(color = "gray90", size = 0.1)
map3<-map2+scale_fill_viridis_c(option = "plasma")
map4<-map3+geom_polygon(data =rus, aes(x = long, y = lat, group = group), colour = "white", fill = "gray85",size=0.2)
map5<-map4+geom_polygon(data =swed, aes(x = long, y = lat, group = group), colour = "white", fill = "gray85",size=0.2)
map6<-map5+geom_polygon(data =nor, aes(x = long, y = lat, group = group), colour = "white", fill = "gray85",size=0.2)
map7<-map6+geom_polygon(data =est, aes(x = long, y = lat, group = group), colour = "white", fill = "gray85",size=0.2)
map8<-map7+geom_text(aes(label="RUSSIA",x=31,y=61),size=4)+geom_text(aes(label="NORWAY",x=24,y=69.5),size=4)+geom_text(aes(label="SWEDEN",x=20,y=67),size=4)
png(file = "finnmap.png", width =2000, height=3000,units='px',res=300)
map8
dev.off()
```


